Performance Models for Data Transfers: A Case Study with Molecular Chemistry Kernels

### 2019-04-20 ICPP-19/communications.tex
attempt to fit the document in 10 pages

	%
	% The abstract is a short summary of the work to be presented in the article.
	\begin{abstract}
		%%		\todo[inline]{Initial submission deadline: April 15}
		With increasing complexity of hardwares, systems with different memory nodes are ubiquitous in High Performance Computing (HPC). It is paramount to develop strategies to overlap the data transfers between memory nodes with computations in order to exploit the full potential of these systems. In this paper, we consider the problem of deciding the order of data transfers between two memory nodes for a set of independent tasks with the objective to minimize the makespan. We prove that with limited memory capacity, obtaining the optimal order of data transfers is a NP-complete problem. We propose several heuristics for this problem and provide details about their favorable situations. We present an analysis of our heuristics on traces, obtained by running 2 molecular chemistry kernels, namely, Hartree–Fock (HF) and Coupled Cluster Single Double (CCSD) on 10 nodes of an HPC system. Our results show that some of our heuristics achieve significant overlap for moderate memory capacities and are very close to the lower bound of makespan.
	\end{abstract}

@@

	Certain applications such as dense linear algebra kernels have regular structure. Therefore, it is possible to associate priorities to computations, based on the task graph structure, and to use them at runtime to make the execution efficient. In irregular applications, programmers do not know the precise structure of the task graphs in advance: tasks are added recursively based on certain sentinel constraints. For such applications, the runtime system sees a set of independent tasks and schedules them on different processing units. It is extremely important for runtimes to decide the order of data transfers for these scheduled computations so as to maximize the overlap between computations and communications. This is the main topic of this paper. We prove that the order of communications on two memory nodes with the objective of minimizing the makespan is a NP-Complete problem if the memory of the target node is limited. Our proof is inspired from work by~\cite{Papadimitriou:1980:FSL:322203.322213}, which applies a similar technique for 2-machine flowshop problem with bounded capacity. The main difference between both approaches is that they consider all tasks have the same occupation on the second machine and the memory occupation starts when the processing finishes on the first machine. On the contrary, our approach is designed for tasks appearing in scientific workloads whose memory requirements are highly irregular and we consider that memory is acquired before starting the data transfer on the communication resource. We propose different runtime strategies in order to maximize the overlap of computations and communications. We evaluate our strategies on the context of a cluster of homogeneous nodes. However, our approach is generic and easily adaptable to any system which operates on different memory spaces. Here are the important contributions of this paper:

@@

	The outline of the paper is the following. Section~\ref{sec:relatedWork} describes past work on the computations with limited memory and similar problems in the literature. In section~\ref{sec:theoreticalProof}, we present an algorithm to obtain the order of data transfers when there is not any memory capacity restriction. Then, we also prove that in general data transfer problem is NP-complete. In Section~\ref{sec:heuristics}, we propose several heuristics and describe their favorable scenarios. We mainly consider three categories of heuristics: static heuristics, dynamic heuristics and static heuristics with dynamic corrections. Sections~\ref{sec:expSetting} describes our experimental setup and we evaluate our proposed strategies on two molecular chemistry kernels in Section~\ref{sec:expResults}. Our results show that static heuristics with dynamic corrections achieve good performance in most cases. We finally propose conclusions and perspectives in Section~\ref{sec:conclusion}.

@@

	The problem of scheduling tasks has been highly studied in the literature and many formulations are known to be NP-Complete~\cite{GareyJohnson}. Many static and dynamic strategies have been proposed and analyzed for scheduling tasks on heterogeneous resources~\cite{heft-Topcuoglu,hipc16multiresource,ipdps16starpu}. There is also a number of studies in the direction of task scheduling with the emphasis on improving locality and minimizing the communication cost~\cite{starpu,heft-Topcuoglu}. Stanisic et. al~\cite{luka-dmdar} proposed a heuristic to schedule tasks on a computational resource where most of its data is available. A similar approach has been adopted by Agullo et. al for the scheduling of sparse linear algebra kernels~\cite{agullo_fmm}. Predari et. al proposed heuristics to partition the task graph across a number of processors such that inter-processor communication can be minimized~\cite{predari:tel-01518956}.

@@

	%%	\todo[inline]{Find a citation saying 3-machine flowshop problem is NP-Complete.}
	%%%%	done:\todo[inline]{Rewrite both paragraphs together.}
	%%	In general, order of task execution with input and output data transfers can be 
	%%	viewed as 3-machine flowshop problem, where processing time on 1st machine is 
	%%	input data transfer time, processing time on 2nd machine is task execution time, 
	%%	processing time on 3rd machine is output data transfer time and the objective is 
	%%	to minimize the total makespan. This is a well known hard problem~\cite{NPComplete3Machine,johnson}. As stated earlier, we do not consider output data separately in our analysis and assume that output data is negligible or stored in a separate buffer. Hence, problem considered in this article is simpler than general 3-machine flowshop problem.

@@

	\begin{algorithm}
		\caption{\label{alg:OrderOfExecutionInfinteMemory}Johnson's~\cite{johnson} algorithm (infinite memory case).}
		\begin{algorithmic}[1]
			\STATE Divide ready tasks in two sets $S_1$ and $S_2$. If computation time of a task $T$ is not less than its communication time, then $T$ is in $S_1$ otherwise in $S_2$.
			\STATE Sort $S_1$ in queue $Q$ by non-decreasing communication times
			\STATE Sort $S_2$ in queue $Q'$ by non-increasing computation times
			\STATE Append $Q'$ to $Q$
			\STATE $\tau_{\text{COMM}} \gets 0$ \hfill\COMMENT{Available time of communication resource}
			\STATE $\tau_{\text{COMP}} \gets 0$\hfill \COMMENT{Available time of computation resource}
			\WHILE{$Q \neq \emptyset$}
			\STATE Remove a task $T$ from beginning of $Q$ for processing
			\STATE $\scomm(T) \gets \tau_{\text{COMM}}$
			\STATE $\scomp(T) \gets max(\scomm(T) + COMM_T, \tau_{\text{COMP}})$
			\STATE $\tau_{\text{COMM}} \gets \scomm(T) + COMM_T$
			\STATE $\tau_{\text{COMP}} \gets \scomp(T) + COMP_T$
			\ENDWHILE
		\end{algorithmic}
	\end{algorithm}

	We also prove optimality of Algorithm~\ref{alg:OrderOfExecutionInfinteMemory} differently in an extended version~\cite{extendedVersion}.
	\subsection{Finite Memory}

@@

			\noindent Memory capacity: $C=b'+3$\\
			\noindent Target makespan: $L=m(b'+3)$
			\caption{Definition of tasks in the reduction from \threepart.}
			\label{table:np.completeness.tasks}
		\end{table}

@@

		\begin{figure}[htb]
			\tikzset{xtick/.style={inner xsep=0pt, inner ysep=3pt, minimum size=0pt, draw},%
				task/.style args={#1start#2length#3res#4color#5}{rounded corners, draw, inner
					sep=0pt, fill=#5, label=center:#1, fit={(#2,#4*0.75) (#2+#3,#4*0.75+0.75)}},%
				vert/.style={inner sep=1pt, fill=black, circle, draw, label=#1}
			}
			\newcommand{\scheduleNoName}[1]{
				\draw[->] (-0.4, 0) -- (#1, 0) node[below] {$t$};
				\draw (0, 0) -- (0, 1.5) node[pos=0.25, left] {Comp.}
				node[pos=0.75, left] {Comm.};
				\draw[dashed,gray] (0, 0.75) -- (#1, 0.75);
			}
			\centering
			\begin{tikzpicture}[yscale=0.7, thick, xscale=0.6]
			\scheduleNoName{12.5}
			\node[task=$A_{1,1}$ start 0 length 1 res 1 color cyan]{};
			\node[task=$A_{1,2}$ start 1 length 1 res 1 color blue!40!white]{};
			\node[task=$A_{1,3}$ start 2 length 1 res 1 color blue!70!white]{};
			\node[task=$K_0$ start 0 length 3 res 0 color gray!40!white]{};
			\node[task=$K_1$ start 3 length 6 res 1 color green]{}; 
			\node[task=$A_{1,1}$ start 3 length 1.8 res 0 color cyan]{};
			\node[task=$A_{1,2}$ start 4.8 length 2.3 res 0 color blue!40!white]{};
			\node[task=$A_{1,3}$ start 7.1 length 1.9 res 0 color blue!70!white]{};
			\node[task=$A_{2,1}$ start 9 length 1 res 1 color cyan]{};
			\node[task=$A_{2,2}$ start 10 length 1 res 1 color blue!40!white]{};
			\node[task=$A_{2,3}$ start 11 length 1 res 1 color blue!70!white]{};
			\node[task=$K_1$ start 9 length 3 res 0 color green]{};
			\draw[<->,thin] (0, -0.2) -- node[below]{$3$} (3, -0.2) ;
			\draw[<->,thin] (9, -0.2) -- node[below]{$3$} (12, -0.2) ;
			\draw[<->,thin] (3, -0.2) -- node[below]{$b'$} (9, -0.2) ;
			\end{tikzpicture}
			\caption{ \label{fig:firstSegment} Pattern of feasible schedule $S$.}
		\end{figure}

@@

		With these $K_i$ tasks in place, the schedule on the computation
		resource contains $m$ slots of length exactly $b'$, in which all $3m$ $A_i$ tasks must fit without preemption. 
		%%\todo[inline]{SKumar: It is not clear why it would be triplet.\\ LED: because two tasks last at most $2\max a'_i =6x$, whereas $b'>6x$. And if any triplet actually contains $4$ tasks, it means one contains only two tasks, and cannot fit in $b'$. Or equivalently, the number of tasks per triplet is at least $3$, and the average number is $3$, hence each triplet contains $3$ tasks.}
		%%	\todo[inline]{At each point in $S$, exactly one $K_i$ task is active, hence we define triplet $TR_{i}$ as the set of tasks which complete their data transfers when $K_i$ task is in computation phase. As $C=b'+3$ and $S$ has no idle time on the communication resource, total computation time of tasks in $TR_{i}$ must not be more than $b'$. $S$ also has no idle time on the computation resource, therefore total computation time of tasks in $TR_i$ is exactly $b'$, and thus $a_{i_1} + a_{i_2} + a_{i_3} = b$. } 
		We thus define $TR_i$ as the set of tasks which execute during the communication phase of task $K_i$.  At each point in $S$, exactly one $K_i$ task is active and the total memory capacity is $b' +3$, hence $TR_i$ contains exactly 3 tasks. Since $S$ has no idle time on the computation resource, the total computation time of tasks in $TR_i$ is exactly $b'$, and thus
		$a_{i_1} + a_{i_2} + a_{i_3} = b$. This partition is thus a valid
		solution for the \threepart instance $A$.

@@

	This theorem shows that adding a memory constraint to our
	problem makes it more difficult, just like it does for limited capacity 2-machine
	flowshop~\cite{Papadimitriou:1980:FSL:322203.322213}. One additional difficulty of our problem however 
	%%       \todo[inline]{I think you mean: infinite memory capacity 2-machine flowshop~\cite{johnson} ? LED: No, I really mean this one. Does this rephrasing help ?} 
	is that it may not be
	optimal to consider the same order on both machines.

@@

	\begin{table}
		\begin{center}
			\begin{tabular}{|c|c|c|c|}
				\hline
				%%				Task & Comm Volume & Comm Time & Comp Time\\ \hline 
				\multirow{2}{*}{Task} & Memory Req & \multirow{2}{*}{Comm Time} & \multirow{2}{*}{Comp Time}\\  
				&=Comm Volume && \\ \hline
				A & 0 & 0 & 5\\ \hline
				B & 4 & 4 & 3\\ \hline
				C & 1 & 1 & 6\\ \hline
				D & 3 & 3 & 7\\ \hline
				E & 6 & 6 & 0.5\\ \hline
				F & 7 & 7 & 0.5\\ \hline
			\end{tabular}
		\end{center}
		\caption{Example instance where ordering on both resources
			has to be different.}
		\label{table:different.order}
	\end{table}

	\begin{figure}
		\centering
		\newcommand{\taskA}[1]{\node[comp=$A$ start #1 length 5 color cyan] {};}
		\newcommand{\taskB}[2][0]{\task[#1]{$B$}{#2}{4}{3}{blue!25!white}}
		\newcommand{\taskC}[2][0]{\task[#1]{$C$}{#2}{1}{6}{blue!50!white}}
		\newcommand{\taskD}[2][0]{\task[#1]{$D$}{#2}{3}{7}{blue!75!white}}
		\newcommand{\taskE}[2][0]{\node[comm=$E$ start #2 length 6 color green!25!white]{};%
			\node[comp={} start #2+6+#1 length 0.5 color green!25!white]{};}
		\newcommand{\taskF}[2][0]{\node[comm=$F$ start #2 length 7 color green!50!white]{};%
			\node[comp={} start #2+7+#1 length 0.5 color green!50!white]{};}

@@

		\subfloat[][Schedule with different ordering on both
		resources]{
			\begin{tikzpicture}[yscale=0.7, thick, xscale=0.3]
			\schedule{24}{}{5,8,14,22}
			\taskA{0}
			\taskB[1]{0}
			\taskC[3]{4}
			\taskD[6.5]{5}
			\taskE{8}
			\taskF{14.5}
			\end{tikzpicture}
			\label{fig:different.order.opt}
		}
		\caption{Schedules for the instance of
			Table~\ref{table:different.order} with a memory capacity of 10.}
		\label{fig:different.order}
	\end{figure}

@@

	Algorithm~\ref{alg:OrderOfExecutionInfinteMemory} presented in Section~\ref{sec:theoreticalProof} achieves an optimal makespan when there is no memory constraint. This optimal value indicates a lower bound on the makespan of the constrained case. We denote this value with \textit{optimal makespan infinite memory} ($OMIM$). In the present Section, we propose different heuristics for the limited memory case, and we assess their efficiency with respect to this lower bound in Section~\ref{sec:expResults}.

	We classify our heuristics into mainly three categories. In the first category, the order of all computations and communications is computed in advance and the same order is followed on both resources. In the second category, the next task to schedule is dyanmically chosen based on different criteria. The final category is based on combining strategies from the first two categories: a static ordering is precomputed and corrected dynamically to avoid idle time caused by memory limitations. In all of our strategies (except linear programming based strategy), communication and computations take place in the same order.

@@

@@

	\end{enumerate}
	\begin{table}[htb]
		\begin{center}

			\begin{tabular}{|c|c|c|c|}
				\hline
				%%				\multirow{2}{*}{Task} & Comm Time & \multirow{2}{*}{Comp Time}\\ 
				%%				&=Comm Volume& \\ \hline
				%%				Task & Comm Volume & Comm Time & Comp Time\\ \hline 
				\multirow{2}{*}{Task} & Memory Req & \multirow{2}{*}{Comm Time} & \multirow{2}{*}{Comp Time}\\  
				&=Comm Volume && \\ \hline
				A & 3 & 3 &  2\\ \hline
				B & 1 & 1 & 3\\ \hline
				C & 4 & 4 & 4\\ \hline
				D & 2 & 2 & 1\\ \hline
			\end{tabular}
			\caption{\label{tab:staticOrderExample} A task set for static order schedules.}
		\end{center}
	\end{table}

	\begin{figure}[htb]
		\newcommand{\taskA}[2][0]{\task[#1]{$A$}{#2}{3}{2}{cyan}}
		\newcommand{\taskB}[2][0]{\task[#1]{$B$}{#2}{1}{3}{blue!40!white}}
		\newcommand{\taskC}[2][0]{\task[#1]{$C$}{#2}{4}{4}{blue!70!white}}
		\newcommand{\taskD}[2][0]{\task[#1]{$D$}{#2}{2}{1}{blue}}

		\centering
		\subfloat[][Infinite Memory Capacity]{
			\begin{tikzpicture}[scale=0.6]
			\schedule{12.5}{OMIM}{1,4,5,8,9,10,11,12}
			\taskB{0}
			\taskC{1}
			\taskA[1]{5}
			\taskD[1]{8}
			\end{tikzpicture}
		}

		\caption{ \label{fig:staticOrderExample} Static order heuristic schedules for Table~\ref{tab:staticOrderExample}.} 
		%%			Optimal schedule length	is 14. 
	\end{figure}

@@

	%% We select a set of tasks which induce minimum idle time on compute resource and satisfy available memory requirement. Then a task is chosen for processing from this set based on the following heuristics.
	\begin{enumerate}[label=\roman*)]
		\item \textit{largest communication task respects memory restriction} ($LCMR$): A task with the largest communication time is chosen. 
		\item \textit{smallest communication task respects memory restriction} ($SCMR$): A task with the smallest communication time is chosen.
		\item \textit{maximum accelerated task respects memory restriction} ($MAMR$): A task with the maximum ratio of computation time to communication time is chosen.
	\end{enumerate}
	\begin{table}[htb]
		\begin{center}

			\begin{tabular}{|c|c|c|c|}
				\hline
				%%				Task & Comm Volume & Comm Time & Comp Time\\ \hline 
				\multirow{2}{*}{Task} & Memory Req & \multirow{2}{*}{Comm Time} & \multirow{2}{*}{Comp Time}\\  
				&=Comm Volume && \\ \hline
				A & 3 & 3 & 2\\ \hline
				B & 1 & 1 &  6\\ \hline
				C & 4 & 4 & 6\\ \hline
				D & 5 & 5 & 1\\ \hline
			\end{tabular}
			\caption{\label{tab:dynamicSelectionExample} A task set for dynamic schedules.}
		\end{center}
	\end{table}

	\begin{figure}[htb]
		\centering

@@

		\centering
		\begin{tikzpicture}[yscale=0.6,xscale=0.3]
		\begin{scope}
		\schedule{24}{LCMR}{1,6,7,8,11,13,17,23}
		\taskB{0}
		\taskD[1]{1}
		\taskA{8}
		\taskC{13}
		\end{scope}
		\begin{scope}[yshift=-2.75cm]
		\schedule{26}{SCMR}{1,4,7,9,13,19,24,25}
		\taskB{0}
		\taskA[3]{1}
		\taskC{9}
		\taskD{19}
		\end{scope}
		\begin{scope}[yshift=-5.5cm]
		\schedule{25}{MAMR}{1,5,7,13,16,18,23,24}
		\taskB{0}
		\taskC[2]{1}
		\taskA{13}
		\taskD{18}
		\end{scope}
		\end{tikzpicture}
		\caption{ \label{fig:dynamicSelectionExample} Different dynamic heuristic schedules for a task set of Table~\ref{tab:dynamicSelectionExample} with a memory capacity of 6.}
		%%		 Optimal schedule length is 23.
	\end{figure}

>>>>

	%
	% The abstract is a short summary of the work to be presented in the article.
	\begin{abstract}
		%%		\todo[inline]{Initial submission deadline: April 15}
		%With increasing hardware complexity and systems with different memory nodes are ubiquitous in High Performance Computing (HPC). 
		In distributed memory systems, it is paramount to develop strategies to overlap the data transfers between memory nodes with the computation in order to exploit the full potential of these systems. In this paper, we consider the problem of determining the order of data transfers between two memory nodes for a set of independent tasks with the objective of minimizing the makespan. We prove that, with limited memory capacity, the problem of obtaining the optimal order of data transfers is NP-complete. We propose several heuristics to determine this order and discuss the conditions that might be favorable to different heuristics. We analyze our heuristics on traces obtained by running 2 molecular chemistry kernels, namely, Hartree–Fock (HF) and Coupled Cluster Single Double (CCSD) on 10 nodes of an HPC system. Our results show that some of our heuristics achieve significant overlap for moderate memory capacities and resulting in makespans that are very close to the lower bound.
	\end{abstract}

@@

	Certain applications such as dense linear algebra kernels have regular structure. Therefore, it is possible to associate priorities to computations, based on the task graph structure, and to use them at runtime to make the execution efficient. In irregular applications, programmers do not know the precise structure of the task graphs in advance: tasks are added recursively based on certain sentinel constraints. For such applications, the runtime system sees a set of independent tasks and schedules them on processing units of different memory spaces. It is extremely important for runtimes to decide the order of data transfers for these scheduled computations so as to maximize the overlap between computations and communications. This is the main topic of this paper. We prove that the order of communications on two memory nodes with the objective of minimizing the makespan is a NP-Complete problem if the memory of the target node is limited. Our proof is inspired from work by~\cite{Papadimitriou:1980:FSL:322203.322213}, which applies a similar technique for 2-machine flowshop problem with bounded capacity. The main difference between both approaches is that they consider all tasks have the same occupation on the second machine and the memory occupation starts when the processing finishes on the first machine. On the contrary, our approach is designed for tasks appearing in scientific workloads whose memory requirements are highly irregular and we consider that memory is acquired before starting the data transfer on the communication resource. We propose different runtime strategies in order to maximize the overlap of computations and communications. We evaluate our strategies on the context of a cluster of homogeneous nodes. However, our approach is generic and easily adaptable to any system which operates on different memory spaces. Here are the important contributions of this paper:

@@

	The outline of the paper is the following. Section~\ref{sec:relatedWork} describes past work on the computations with limited memory and similar problems in the literature. In section~\ref{sec:theoreticalProof}, we present an algorithm to obtain the order of data transfers when there is not any memory capacity restriction. Then, we also prove that in general data transfer problem is NP-complete. In Section~\ref{sec:heuristics}, we propose several heuristics and discuss the conditions that might be favorable to them. We mainly consider three categories of heuristics: static heuristics, dynamic heuristics and static heuristics with dynamic corrections. Sections~\ref{sec:expSetting} describes our experimental setup and we evaluate our proposed strategies on two molecular chemistry kernels in Section~\ref{sec:expResults}. Our results show that static heuristics with dynamic corrections achieve good performance in most cases. We finally propose conclusions and perspectives in Section~\ref{sec:conclusion}.

@@

	The problem of scheduling tasks has been highly studied in the literature and many formulations are known to be NP-Complete~\cite{GareyJohnson}. Many static and dynamic strategies have been proposed and analyzed for scheduling tasks on heterogeneous resources~\cite{heft-Topcuoglu,ipdps16starpu}. There is also a number of studies in the direction of task scheduling with the emphasis on improving locality and minimizing the communication cost~\cite{starpu,heft-Topcuoglu}. Stanisic et. al~\cite{luka-dmdar} proposed a heuristic to schedule tasks on a computational resource where most of its data is available. A similar approach has been adopted by Agullo et. al for the scheduling of sparse linear algebra kernels~\cite{agullo_fmm}. Predari et. al proposed heuristics to partition the task graph across a number of processors such that inter-processor communication can be minimized~\cite{predari:tel-01518956}.

@@

@@

	\begin{algorithm}
		\caption{\label{alg:OrderOfExecutionInfinteMemory}Johnson's~\cite{johnson} algorithm (infinite memory case).}
		\begin{algorithmic}[1]
			\STATE Divide ready tasks in two sets $S_1$ and $S_2$. If computation time of a task $T$ is not less than its communication time, then $T$ is in $S_1$ otherwise in $S_2$.
			\STATE Sort $S_1$ in queue $Q$ by non-decreasing communication times
			\STATE Sort $S_2$ in queue $Q'$ by non-increasing computation times
			\STATE Append $Q'$ to $Q$
			\STATE $\tau_{\text{COMM}} \gets 0$ \hfill\COMMENT{Available time of communication resource}
			\STATE $\tau_{\text{COMP}} \gets 0$\hfill \COMMENT{Available time of computation resource}
			\WHILE{$Q \neq \emptyset$}
			\STATE Remove a task $T$ from beginning of $Q$ for processing
			\STATE $\scomm(T) \gets \tau_{\text{COMM}}$
			\STATE $\scomp(T) \gets max(\scomm(T) + COMM_T, \tau_{\text{COMP}})$
			\STATE $\tau_{\text{COMM}} \gets \scomm(T) + COMM_T$
			\STATE $\tau_{\text{COMP}} \gets \scomp(T) + COMP_T$
			\ENDWHILE
		\end{algorithmic}
	\end{algorithm}
	\vspace*{-0.35cm}
	We also prove optimality of Algorithm~\ref{alg:OrderOfExecutionInfinteMemory} differently in an extended version~\cite{extendedVersion}.
	\subsection{Finite Memory}

@@

			\noindent Memory capacity: $C=b'+3$\\
			\noindent Target makespan: $L=m(b'+3)$
			\vspace*{-0.25cm}\caption{Definition of tasks in the reduction from \threepart.\vspace*{-0.5cm}}
			\label{table:np.completeness.tasks}
		\end{table}

@@

		\begin{figure}[htb]
			\tikzset{xtick/.style={inner xsep=0pt, inner ysep=3pt, minimum size=0pt, draw},%
				task/.style args={#1start#2length#3res#4color#5}{rounded corners, draw, inner
					sep=0pt, fill=#5, label=center:#1, fit={(#2,#4*0.75) (#2+#3,#4*0.75+0.75)}},%
				vert/.style={inner sep=1pt, fill=black, circle, draw, label=#1}
			}
			\newcommand{\scheduleNoName}[1]{
				\draw[->] (-0.4, 0) -- (#1, 0) node[below] {$t$};
				\draw (0, 0) -- (0, 1.5) node[pos=0.25, left] {Comp.}
				node[pos=0.75, left] {Comm.};
				\draw[dashed,gray] (0, 0.75) -- (#1, 0.75);
			}
			\centering
			\begin{tikzpicture}[yscale=0.7, thick, xscale=0.6]
			\scheduleNoName{12.5}
			\node[task=$A_{1,1}$ start 0 length 1 res 1 color cyan]{};
			\node[task=$A_{1,2}$ start 1 length 1 res 1 color blue!40!white]{};
			\node[task=$A_{1,3}$ start 2 length 1 res 1 color blue!70!white]{};
			\node[task=$K_0$ start 0 length 3 res 0 color gray!40!white]{};
			\node[task=$K_1$ start 3 length 6 res 1 color green]{}; 
			\node[task=$A_{1,1}$ start 3 length 1.8 res 0 color cyan]{};
			\node[task=$A_{1,2}$ start 4.8 length 2.3 res 0 color blue!40!white]{};
			\node[task=$A_{1,3}$ start 7.1 length 1.9 res 0 color blue!70!white]{};
			\node[task=$A_{2,1}$ start 9 length 1 res 1 color cyan]{};
			\node[task=$A_{2,2}$ start 10 length 1 res 1 color blue!40!white]{};
			\node[task=$A_{2,3}$ start 11 length 1 res 1 color blue!70!white]{};
			\node[task=$K_1$ start 9 length 3 res 0 color green]{};
			\draw[<->,thin] (0, -0.2) -- node[below]{$3$} (3, -0.2) ;
			\draw[<->,thin] (9, -0.2) -- node[below]{$3$} (12, -0.2) ;
			\draw[<->,thin] (3, -0.2) -- node[below]{$b'$} (9, -0.2) ;
			\end{tikzpicture}
			\caption{ \label{fig:firstSegment} Pattern of feasible schedule $S$.\vspace*{-0.3cm}}
		\end{figure}

@@

		With these $K_i$ tasks in place, the schedule on the computation
		resource contains $m$ slots of length exactly $b'$, in which all $3m$ $A_i$ tasks must fit without preemption. 
		%%\todo[inline]{SKumar: It is not clear why it would be triplet.\\ LED: because two tasks last at most $2\max a'_i =6x$, whereas $b'>6x$. And if any triplet actually contains $4$ tasks, it means one contains only two tasks, and cannot fit in $b'$. Or equivalently, the number of tasks per triplet is at least $3$, and the average number is $3$, hence each triplet contains $3$ tasks.}
		%%	\todo[inline]{At each point in $S$, exactly one $K_i$ task is active, hence we define triplet $TR_{i}$ as the set of tasks which complete their data transfers when $K_i$ task is in computation phase. As $C=b'+3$ and $S$ has no idle time on the communication resource, total computation time of tasks in $TR_{i}$ must not be more than $b'$. $S$ also has no idle time on the computation resource, therefore total computation time of tasks in $TR_i$ is exactly $b'$, and thus $a_{i_1} + a_{i_2} + a_{i_3} = b$. } 
		We thus define $TR_i$ as the set of tasks which execute during the communication phase of task $K_i$.  At each point in $S$, exactly one $K_i$ task is active and the total memory capacity is $b' +3$, hence $TR_i$ contains exactly 3 $A_i$ tasks. Since $S$ has no idle time on the computation resource, the total computation time of tasks in $TR_i$ is exactly $b'$, and thus
		$a_{i_1} + a_{i_2} + a_{i_3} = b$. This partition is thus a valid
		solution for the \threepart instance $A$.

@@

	This theorem shows that adding a memory constraint to our
	problem makes it more difficult, just like it does for limited capacity 2-machine
	flowshop~\cite{Papadimitriou:1980:FSL:322203.322213}. One additional difficulty of our problem however is that it may not be
	optimal to consider the same order on both machines.

@@

	\begin{table}[!ht]
		\begin{center}
			\begin{tabular}{|c|c|c|c|}
				\hline
				%%				Task & Comm Volume & Comm Time & Comp Time\\ \hline 
				\multirow{2}{*}{Task} & Memory Req & \multirow{2}{*}{Comm Time} & \multirow{2}{*}{Comp Time}\\  
				&=Comm Volume && \\ \hline
				A & 0 & 0 & 5\\ \hline
				B & 4 & 4 & 3\\ \hline
				C & 1 & 1 & 6\\ \hline
				D & 3 & 3 & 7\\ \hline
				E & 6 & 6 & 0.5\\ \hline
				F & 7 & 7 & 0.5\\ \hline
			\end{tabular}
		\end{center}
		\caption{Example instance where ordering on both resources
			has to be different\vspace*{-0.9cm}.}
		\label{table:different.order}
	\end{table}

	\begin{figure}[!htb]
		\centering
		\newcommand{\taskA}[1]{\node[comp=$A$ start #1 length 5 color cyan] {};}
		\newcommand{\taskB}[2][0]{\task[#1]{$B$}{#2}{4}{3}{blue!25!white}}
		\newcommand{\taskC}[2][0]{\task[#1]{$C$}{#2}{1}{6}{blue!50!white}}
		\newcommand{\taskD}[2][0]{\task[#1]{$D$}{#2}{3}{7}{blue!75!white}}
		\newcommand{\taskE}[2][0]{\node[comm=$E$ start #2 length 6 color green!25!white]{};%
			\node[comp={} start #2+6+#1 length 0.5 color green!25!white]{};}
		\newcommand{\taskF}[2][0]{\node[comm=$F$ start #2 length 7 color green!50!white]{};%
			\node[comp={} start #2+7+#1 length 0.5 color green!50!white]{};}

@@

		\subfloat[][Schedule with different ordering on both
		resources]{
			\begin{tikzpicture}[yscale=0.7, thick, xscale=0.3]
			\schedule{24}{}{5,8,14,22}
			\taskA{0}
			\taskB[1]{0}
			\taskC[3]{4}
			\taskD[6.5]{5}
			\taskE{8}
			\taskF{14.5}
			\end{tikzpicture}
			\label{fig:different.order.opt}
		}
		\caption{Schedules for the instance of
			Table~\ref{table:different.order} with a memory capacity of 10.\vspace*{-0.75cm}}
		\label{fig:different.order}
	\end{figure}

@@

	Algorithm~\ref{alg:OrderOfExecutionInfinteMemory} presented in Section~\ref{sec:theoreticalProof} achieves an optimal makespan when there is no memory constraint. This optimal value indicates a lower bound on the makespan of the constrained case. We denote this value with \textit{optimal makespan infinite memory} ($OMIM$). In this section, we propose different heuristics for the limited memory case, and we assess their efficiency with respect to $OMIM$ in Section~\ref{sec:expResults}.

	We classify our heuristics into mainly three categories. In the first category, the order of all computations and communications is computed in advance and the same order is followed on both resources. In the second category, the next task to schedule is dynamically chosen based on different criteria. The final category is based on combining strategies from the first two categories: a static ordering is precomputed and corrected dynamically to avoid idle time caused by memory limitations. In all of our strategies (except linear programming based strategy), communication and computations take place in the same order.

@@

\begin{figure}[!tb]
	\newcommand{\taskA}[2][0]{\task[#1]{$A$}{#2}{3}{2}{cyan}}
	\newcommand{\taskB}[2][0]{\task[#1]{$B$}{#2}{1}{3}{blue!40!white}}
	\newcommand{\taskC}[2][0]{\task[#1]{$C$}{#2}{4}{4}{blue!70!white}}
	\newcommand{\taskD}[2][0]{\task[#1]{$D$}{#2}{2}{1}{blue}}
\vspace*{-0.25cm}	
	\centering
	\subfloat[][Infinite Memory Capacity]{
		\begin{tikzpicture}[scale=0.6]
		\schedule{12.5}{OMIM}{1,4,5,8,9,10,11,12}
		\taskB{0}
		\taskC{1}
		\taskA[1]{5}
		\taskD[1]{8}
		\end{tikzpicture}
	}

	\caption{\vspace*{-0.2cm} \label{fig:staticOrderExample} Static order heuristic schedules for Table~\ref{tab:staticOrderExample}.} 
	%%			Optimal schedule length	is 14. 
\end{figure}

	\begin{table}[!tb]
	\begin{center}
%%\vspace*{-0.10cm}		
		\begin{tabular}{|c|c|c|c|}
			\hline
			%%				\multirow{2}{*}{Task} & Comm Time & \multirow{2}{*}{Comp Time}\\ 
			%%				&=Comm Volume& \\ \hline
			%%				Task & Comm Volume & Comm Time & Comp Time\\ \hline 
			\multirow{2}{*}{Task} & Memory Req & \multirow{2}{*}{Comm Time} & \multirow{2}{*}{Comp Time}\\  
			&=Comm Volume && \\ \hline
			A & 3 & 3 &  2\\ \hline
			B & 1 & 1 & 3\\ \hline
			C & 4 & 4 & 4\\ \hline
			D & 2 & 2 & 1\\ \hline
		\end{tabular}
		\caption{\label{tab:staticOrderExample} A task set for static order schedules.\vspace*{-0.75cm}}
	\end{center}
\end{table}

@@

	\end{enumerate}

@@

	%% We select a set of tasks which induce minimum idle time on compute resource and satisfy available memory requirement. Then a task is chosen for processing from this set based on the following heuristics.
	\begin{enumerate}[label=\roman*)]
		\item \textit{largest communication task respects memory restriction} ($LCMR$): A task with the largest communication time is chosen. 
		\item \textit{smallest communication task respects memory restriction} ($SCMR$): A task with the smallest communication time is chosen.
		\item \textit{maximum accelerated task respects memory restriction} ($MAMR$): A task with the maximum ratio of computation time to communication time is chosen.
	\end{enumerate}

	\begin{figure}[!t]
		\centering

@@

		\centering
		\begin{tikzpicture}[yscale=0.6,xscale=0.3]
		\begin{scope}
		\schedule{24}{LCMR}{1,6,7,8,11,13,17,23}
		\taskB{0}
		\taskD[1]{1}
		\taskA{8}
		\taskC{13}
		\end{scope}
		\begin{scope}[yshift=-2.75cm]
		\schedule{26}{SCMR}{1,4,7,9,13,19,24,25}
		\taskB{0}
		\taskA[3]{1}
		\taskC{9}
		\taskD{19}
		\end{scope}
		\begin{scope}[yshift=-5.5cm]
		\schedule{25}{MAMR}{1,5,7,13,16,18,23,24}
		\taskB{0}
		\taskC[2]{1}
		\taskA{13}
		\taskD{18}
		\end{scope}
		\end{tikzpicture}
		\caption{ \label{fig:dynamicSelectionExample} Different dynamic heuristic schedules for a task set of Table~\ref{tab:dynamicSelectionExample} with a memory capacity of 6.\vspace*{-0.25cm}}
		%%		 Optimal schedule length is 23.
	\end{figure}

	\begin{table}[!tb]
	\begin{center}
		\begin{tabular}{|c|c|c|c|}
			\hline
			%%				Task & Comm Volume & Comm Time & Comp Time\\ \hline 
			\multirow{2}{*}{Task} & Memory Req & \multirow{2}{*}{Comm Time} & \multirow{2}{*}{Comp Time}\\  
			&=Comm Volume && \\ \hline
			A & 3 & 3 & 2\\ \hline
			B & 1 & 1 &  6\\ \hline
			C & 4 & 4 & 6\\ \hline
			D & 5 & 5 & 1\\ \hline
		\end{tabular}
		\caption{\label{tab:dynamicSelectionExample} A task set for dynamic schedules.\vspace*{-0.5cm}}
	\end{center}
	\end{table}
